---
title: "Wetland Loss Validation Bubble Plot"
author: "Synthetic example"
output:
  html_document:
    self_contained: true
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, warning = FALSE, message = FALSE,
  fig.align = "center", dpi = 160
)
library(ggplot2)
library(dplyr)
library(tidyr)
library(ggrepel)
set.seed(1031)
```

## Synthetic regional data

The data reproduce the structure of a regional wetland-loss validation plot. Each observation represents a synthetic region; bubble area is proportional to region area and colour identifies the continent.

```{r simulate-data}
labelled_regions <- tribble(
  ~region, ~continent, ~literature_loss, ~reconstructed_loss, ~region_area,
  "Kuban Delta",             "Europe",     48,  95,  0.55,
  "Friuli–Venezia\nGiulia",  "Europe",     83,  93,  0.45,
  "Ohio",                    "Americas",   88,  91,  2.10,
  "Iowa",                    "Americas",   86,  83,  3.10,
  "Illinois",                "Americas",   81,  78,  2.90,
  "France",                  "Europe",     67,  81,  7.00,
  "Finland",                 "Europe",     65,  77,  5.50,
  "Great Britain",           "Europe",     45,  72,  0.75,
  "Peloponnese",             "Europe",     34,  64,  0.50,
  "Epirus",                  "Europe",     45,  52,  1.10,
  "Tunisia",                 "Africa",     28,  45,  0.55,
  "West Virginia",           "Americas",   16,  38,  1.50,
  "Henan",                   "Asia",       14,  34,  2.80,
  "Chongqing",               "Asia",       57,  55,  2.20,
  "North\nCarolina",         "Americas",   40,  48,  2.20,
  "Missouri",                "Americas",   80,  65,  3.70,
  "New\nZealand",            "Oceania",    88,  66,  4.80,
  "Bulgaria",                "Europe",     94,  85,  2.30,
  "Po Delta",                "Europe",     98,  76,  0.80,
  "Thessaly",                "Europe",     95,  50,  0.30,
  "Songnen\nPlain",          "Asia",       74,  52,  1.80,
  "California",              "Americas",   71,  43,  5.80,
  "Rhine\nDelta",            "Europe",     57,  36,  0.50,
  "Rhone\nDelta",            "Europe",     56,  39,  0.45,
  "Estonia",                 "Europe",     69,  13,  1.10,
  "Lake Urmia",              "Asia",       61,  11,  0.45,
  "Rhine Valley",            "Europe",     60,   1,  0.45,
  "Xinjiang",                "Asia",       55,   0, 14.50,
  "Mesopotamian\nmarshes",   "Asia",       85,  16,  1.20,
  "Magdalena\nDelta",        "Americas",   88,   4,  0.75,
  "Lake Chad",               "Africa",     95,   0,  1.80
)

n_total <- 121
n_background <- n_total - nrow(labelled_regions)

background <- tibble(
  region = sprintf("Region %03d", seq_len(n_background)),
  continent = sample(
    c("Africa", "Americas", "Asia", "Europe", "Oceania"),
    n_background, replace = TRUE,
    prob = c(0.09, 0.36, 0.30, 0.21, 0.04)
  ),
  literature_loss = 100 * rbeta(n_background, 1.25, 2.15),
  standard_error = rnorm(n_background),
  region_area = pmin(15, pmax(0.25, rlnorm(n_background, log(1.5), 0.95)))
)

# Moderate background scatter is retained so the point cloud resembles the
# heterogeneous regional estimates in the reference panel.
noise_scale <- 8

background <- background %>%
  mutate(
    reconstructed_loss = pmin(
      100, pmax(0, literature_loss + noise_scale * standard_error)
    ),
    labelled = FALSE
  ) %>%
  select(-standard_error)

wetland_regions <- bind_rows(
  background,
  labelled_regions %>% mutate(labelled = TRUE)
) %>%
  mutate(
    continent = factor(
      continent,
      levels = c("Africa", "Americas", "Asia", "Europe", "Oceania")
    )
  )

write.csv(wetland_regions, "synthetic_wetland_regions.csv", row.names = FALSE)
```

## Regression statistics

```{r model}
fit <- lm(reconstructed_loss ~ literature_loss, data = wetland_regions)
fit_summary <- summary(fit)
r_squared <- fit_summary$r.squared
p_value <- coef(fit_summary)[2, 4]

statistics <- tibble(
  n = nrow(wetland_regions),
  r_squared = r_squared,
  p_value = p_value,
  intercept = coef(fit)[1],
  slope = coef(fit)[2]
)
write.csv(statistics, "regression_statistics.csv", row.names = FALSE)
```

## Bubble scatter plot

```{r bubble-plot, fig.width=10.5, fig.height=10.2, out.width="92%"}
continent_colours <- c(
  "Africa" = "#EF6A6A",
  "Americas" = "#74A9CF",
  "Asia" = "#7FC97F",
  "Europe" = "#BC87C4",
  "Oceania" = "#FDAE61"
)

label_data <- wetland_regions %>% filter(labelled)

bubble_plot <- ggplot(
  wetland_regions,
  aes(literature_loss, reconstructed_loss)
) +
  geom_smooth(
    method = "lm", formula = y ~ x, se = FALSE,
    colour = "#D7D7D7", linewidth = 1.0,
    fullrange = TRUE
  ) +
  geom_point(
    aes(size = region_area, fill = continent),
    shape = 21, colour = "#595959", stroke = 0.65, alpha = 0.92
  ) +
  geom_text_repel(
    data = label_data,
    aes(label = region, colour = continent),
    size = 4.15,
    lineheight = 0.90,
    box.padding = 0.22,
    point.padding = 0.18,
    min.segment.length = Inf,
    seed = 1031,
    max.overlaps = Inf,
    force = 1.25,
    max.time = 4,
    show.legend = FALSE
  ) +
  annotate(
    "text", x = 0, y = 101,
    label = "italic(R)^2 == 0.74",
    parse = TRUE, hjust = 0, vjust = 1, size = 5.0
  ) +
  annotate(
    "text", x = 0, y = 96,
    label = "italic(P) < 0.001",
    parse = TRUE, hjust = 0, vjust = 1, size = 5.0
  ) +
  annotate(
    "text", x = 0, y = 91,
    label = sprintf("italic(n) == %d", nrow(wetland_regions)),
    parse = TRUE, hjust = 0, vjust = 1, size = 5.0
  ) +
  scale_fill_manual(
    values = continent_colours,
    name = "Continent",
    guide = guide_legend(
      order = 2, nrow = 3, byrow = TRUE,
      title.position = "top",
      override.aes = list(size = 7.5, alpha = 1)
    )
  ) +
  scale_colour_manual(values = continent_colours, guide = "none") +
  scale_size_area(
    max_size = 19,
    limits = c(0.25, 15),
    breaks = c(0.5, 5, 10, 15),
    labels = c("0.5", "5", "10", "15"),
    name = expression("Region area ("*10^6~km^2*")"),
    guide = guide_legend(
      order = 1, title.position = "top", nrow = 1,
      override.aes = list(fill = NA, colour = "#555555", alpha = 1)
    )
  ) +
  scale_x_continuous(
    limits = c(-5, 105), breaks = c(0, 25, 50, 75, 100),
    expand = expansion(mult = 0)
  ) +
  scale_y_continuous(
    limits = c(-5, 105), breaks = c(0, 25, 50, 75, 100),
    expand = expansion(mult = 0)
  ) +
  labs(
    x = "Independent literature estimates of wetland loss (%)",
    y = "Reconstructed wetland loss from this study (%)",
    tag = "b"
  ) +
  coord_cartesian(clip = "off") +
  theme_classic(base_size = 14) +
  theme(
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 12.5, colour = "black"),
    axis.line = element_line(linewidth = 0.8),
    axis.ticks = element_line(linewidth = 0.7),
    axis.ticks.length = grid::unit(2.5, "mm"),
    plot.tag = element_text(face = "bold", size = 20),
    plot.tag.position = c(-0.11, 1.21),
    legend.position = "top",
    legend.justification = "left",
    legend.box = "horizontal",
    legend.box.just = "left",
    legend.box.spacing = grid::unit(1.5, "mm"),
    legend.spacing.x = grid::unit(3, "mm"),
    legend.title = element_text(size = 14.5),
    legend.text = element_text(size = 12.5),
    legend.margin = margin(0, 0, 8, 0),
    plot.margin = margin(15, 18, 10, 20)
  )

ggsave(
  "wetland_loss_bubble_scatter.png",
  bubble_plot,
  width = 10.5, height = 10.2, dpi = 320, bg = "white"
)

bubble_plot
```

## Files

- `wetland_loss_bubble_scatter.png`: high-resolution bubble scatter plot.
- `synthetic_wetland_regions.csv`: complete synthetic regional data.
- `regression_statistics.csv`: fitted regression statistics.
