---
title: "Three-dimensional Gibbs Energy Landscape"
author: "Synthetic example"
output:
  html_document:
    self_contained: true
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.align = "center", dpi = 160)
library(dplyr)
library(tidyr)
library(plot3D)
set.seed(1031)
```

## Model

The surface uses a synthetic Gaussian-mixture probability distribution. Relative Gibbs energy is calculated as

\[
\Delta G(x,y)=\min\left[-1.45\log\{P(x,y)/P_{\max}\}, 11\right].
\]

```{r simulate-landscape}
x_values <- seq(0.020, 0.190, length.out = 170)
y_values <- seq(1.458, 1.535, length.out = 170)
landscape <- expand_grid(PC1 = x_values, PC2 = y_values)

gaussian2d <- function(x, y, mux, muy, sdx, sdy, rho = 0) {
  zx <- (x - mux) / sdx
  zy <- (y - muy) / sdy
  exp(-(zx^2 - 2 * rho * zx * zy + zy^2) / (2 * (1 - rho^2))) /
    (2 * pi * sdx * sdy * sqrt(1 - rho^2))
}

states <- tribble(
  ~weight, ~mux,  ~muy,  ~sdx,  ~sdy,  ~rho,
     0.28, 0.057, 1.471, 0.009, 0.0055,  0.25,
     0.07, 0.076, 1.481, 0.010, 0.0060,  0.45,
     0.02, 0.097, 1.489, 0.011, 0.0065,  0.30,
     0.21, 0.128, 1.478, 0.009, 0.0065,  0.15,
     0.27, 0.144, 1.494, 0.011, 0.0100,  0.55,
     0.15, 0.159, 1.512, 0.010, 0.0085,  0.48
)

density <- rep(0, nrow(landscape))
for (i in seq_len(nrow(states))) {
  density <- density + states$weight[i] * gaussian2d(
    landscape$PC1, landscape$PC2,
    states$mux[i], states$muy[i],
    states$sdx[i], states$sdy[i], states$rho[i]
  )
}

modulation <- 1 + 0.045 * sin(165 * landscape$PC1) *
  cos(310 * landscape$PC2)
density <- pmax(density * modulation, .Machine$double.xmin)

landscape <- landscape %>%
  mutate(
    probability = density / max(density),
    gibbs_energy = pmin(1.45 * -log(probability), 11)
  )

z_matrix <- xtabs(gibbs_energy ~ PC1 + PC2, data = landscape)

write.csv(landscape, "gibbs_energy_grid_3d.csv", row.names = FALSE)
write.csv(states, "synthetic_state_parameters.csv", row.names = FALSE)
```

## Three-dimensional surface

The coloured projection on the lower plane preserves the two-dimensional landscape while the surface displays the energy barriers in three dimensions.

```{r surface, fig.width=9.0, fig.height=7.5, out.width="95%"}
energy_colours <- colorRampPalette(c(
  "#071B8C", "#003BCE", "#0078F0", "#00C8F0",
  "#35E7B0", "#A9F06B", "#FFF33B", "#FF9B24",
  "#EE321A", "#9E0000"
))(180)

draw_landscape_3d <- function() {
  par(
    mar = c(3.5, 3.2, 3.2, 5.6),
    mgp = c(1.8, 0.55, 0),
    tcl = -0.25,
    cex.axis = 0.72,
    cex.lab = 0.88,
    family = "sans"
  )

  persp3D(
    x = x_values,
    y = y_values,
    z = z_matrix,
    colvar = z_matrix,
    col = energy_colours,
    clim = c(0, 11),
    zlim = c(0, 11.3),
    phi = 24,
    theta = 38,
    expand = 0.72,
    border = NA,
    facets = TRUE,
    shade = 0.08,
    lighting = FALSE,
    image = TRUE,
    contour = list(
      side = "zmin", levels = seq(0, 10, by = 2),
      col = "#F5F5F5", lwd = 0.7
    ),
    bty = "b2",
    ticktype = "detailed",
    xlab = "PC1",
    ylab = "PC2",
    zlab = "Gibbs free energy (kcal/mol)",
    main = "Three-dimensional Gibbs Energy Landscape",
    clab = "Gibbs free energy (kcal/mol)",
    colkey = list(
      side = 4, length = 0.58, width = 0.65,
      shift = 0.05, cex.axis = 0.72, cex.clab = 0.78
    )
  )

  mtext("H", side = 3, line = 1.05, adj = 0.01, cex = 2.0, font = 2)
}

png(
  "gibbs_energy_landscape_3d.png",
  width = 2880, height = 2400, res = 320,
  bg = "white", type = "cairo"
)
draw_landscape_3d()
dev.off()

draw_landscape_3d()
```

## Files

- `gibbs_energy_landscape_3d.png`: high-resolution 3D surface.
- `gibbs_energy_grid_3d.csv`: complete energy grid.
- `synthetic_state_parameters.csv`: Gaussian-state parameters.
