---
title: "Two-dimensional Gibbs Energy Landscape"
author: "Synthetic example"
output:
  html_document:
    self_contained: true
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.align = "center", dpi = 160)
library(ggplot2)
library(dplyr)
library(tidyr)
set.seed(1031)
```

## Model

The landscape is calculated from a synthetic mixture of correlated Gaussian conformational states. The relative Gibbs energy is defined as

\[
\Delta G(x,y)=-\log\{P(x,y)/P_{\max}\},
\]

and truncated at 11 kcal/mol to match the finite colour scale.

```{r simulate-landscape}
x_values <- seq(0.020, 0.190, length.out = 220)
y_values <- seq(1.458, 1.535, length.out = 220)
landscape <- expand_grid(PC1 = x_values, PC2 = y_values)

gaussian2d <- function(x, y, mux, muy, sdx, sdy, rho = 0) {
  zx <- (x - mux) / sdx
  zy <- (y - muy) / sdy
  exp(-(zx^2 - 2 * rho * zx * zy + zy^2) / (2 * (1 - rho^2))) /
    (2 * pi * sdx * sdy * sqrt(1 - rho^2))
}

states <- tribble(
  ~weight, ~mux,  ~muy,  ~sdx,  ~sdy,  ~rho,
     0.28, 0.057, 1.471, 0.009, 0.0055,  0.25,
     0.07, 0.076, 1.481, 0.010, 0.0060,  0.45,
     0.02, 0.097, 1.489, 0.011, 0.0065,  0.30,
     0.21, 0.128, 1.478, 0.009, 0.0065,  0.15,
     0.27, 0.144, 1.494, 0.011, 0.0100,  0.55,
     0.15, 0.159, 1.512, 0.010, 0.0085,  0.48
)

density <- rep(0, nrow(landscape))
for (i in seq_len(nrow(states))) {
  density <- density + states$weight[i] * gaussian2d(
    landscape$PC1, landscape$PC2,
    states$mux[i], states$muy[i],
    states$sdx[i], states$sdy[i], states$rho[i]
  )
}

# Small deterministic modulation gives the basin boundaries an organic shape.
modulation <- 1 + 0.045 * sin(165 * landscape$PC1) *
  cos(310 * landscape$PC2)
density <- pmax(density * modulation, .Machine$double.xmin)

landscape <- landscape %>%
  mutate(
    probability = density / max(density),
    gibbs_energy = pmin(1.45 * -log(probability), 11)
  )

write.csv(landscape, "gibbs_energy_grid_2d.csv", row.names = FALSE)
write.csv(states, "synthetic_state_parameters.csv", row.names = FALSE)
```

## Gibbs energy heatmap

```{r heatmap, fig.width=7.2, fig.height=6.6, out.width="85%"}
energy_colours <- c(
  "#071B8C", "#003BCE", "#0078F0", "#00C8F0",
  "#35E7B0", "#A9F06B", "#FFF33B", "#FF9B24",
  "#EE321A", "#9E0000"
)

p2d <- ggplot(landscape, aes(PC1, PC2, fill = gibbs_energy)) +
  geom_raster(interpolate = TRUE) +
  geom_contour(
    aes(z = gibbs_energy),
    breaks = c(2, 4, 6, 8),
    colour = "white", linewidth = 0.18, alpha = 0.34
  ) +
  scale_fill_gradientn(
    colours = energy_colours,
    limits = c(0, 11), breaks = seq(0, 10, 2),
    name = expression(Delta*G~"(kcal/mol)"),
    guide = guide_colourbar(
      title.position = "right", title.theme = element_text(angle = 90),
      barheight = grid::unit(52, "mm"), barwidth = grid::unit(5, "mm")
    )
  ) +
  scale_x_continuous(expand = c(0, 0), breaks = seq(0.02, 0.18, 0.04)) +
  scale_y_continuous(expand = c(0, 0), breaks = c(1.46, 1.48, 1.50, 1.52)) +
  labs(
    title = "Gibbs Energy Landscape",
    subtitle = "Two metastable basins in PC space",
    x = "PC1", y = "PC2",
    tag = "G"
  ) +
  theme_classic(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(size = 10, hjust = 0.5, colour = "grey35"),
    plot.tag = element_text(face = "bold", size = 24),
    plot.tag.position = c(0.02, 0.98),
    axis.title = element_text(face = "bold"),
    axis.line = element_line(linewidth = 0.65),
    axis.ticks = element_line(linewidth = 0.55),
    legend.title = element_text(size = 10),
    plot.margin = margin(12, 12, 8, 12)
  )

ggsave(
  "gibbs_energy_landscape_2d.png", p2d,
  width = 7.2, height = 6.6, dpi = 320, bg = "white"
)
p2d
```

## Files

- `gibbs_energy_landscape_2d.png`: high-resolution heatmap.
- `gibbs_energy_grid_2d.csv`: complete energy grid.
- `synthetic_state_parameters.csv`: Gaussian-state parameters.
