---
title: "Annotated Correlation Bubble Heatmap"
author: "Synthetic example"
output:
  html_document:
    self_contained: true
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  dpi = 160
)

library(ggplot2)
library(dplyr)
library(tidyr)
library(patchwork)

set.seed(1031)
```

## Overview

This example recreates a Cell-style annotated correlation matrix using fully synthetic data. Circle colour represents the signed Spearman correlation, while circle area represents its absolute magnitude. The annotation strips describe each signature's biological category, source context, and cell-cycle enrichment. PSA and Ki67 are displayed as auxiliary correlations at the right and bottom edges.

## Simulate signature scores

```{r simulate-data}
signature_groups <- list(
  `AR/SR signaling` = c(
    "AR_Spratt", "AR_Hieronymus", "AR_Full", "AR_V7",
    "Androgen_Response", "Basal_Zhang"
  ),
  `DNA repair` = c(
    "BRCA1_DDR", "Homologous_Recomb", "DNA_Repair", "MMR",
    "Chromatin_Remodel", "Replication_Stress"
  ),
  `Immune / microenvironment` = c(
    "CD8_Tcell", "CD4_Tcell", "Bcell", "IFN_Gamma",
    "Inflammatory", "TGFbeta", "Stromal"
  ),
  `Lineage / plasticity` = c(
    "Luminal_A", "Luminal_B", "Basal", "Neuroendocrine",
    "EMT", "Stemness"
  ),
  `Oncogenic pathways` = c(
    "PI3K_AKT", "MAPK", "MYC", "WNT", "TP53_Loss", "RB_Loss"
  ),
  `Metabolism / prognosis` = c(
    "Hypoxia", "Glycolysis", "OxPhos", "Proliferation",
    "Radiotherapy", "Metastasis"
  )
)

signature_meta <- tibble(
  signature = unlist(signature_groups, use.names = FALSE),
  biological_category = rep(names(signature_groups), lengths(signature_groups))
) %>%
  mutate(
    origin = sample(
      c("Prostate cancer", "Pan cancer", "CRPC", "Breast", "Melanoma"),
      n(), replace = TRUE,
      prob = c(0.42, 0.24, 0.16, 0.10, 0.08)
    ),
    cell_cycle_enriched = if_else(
      signature %in% c("Replication_Stress", "MYC", "RB_Loss", "Proliferation", "Radiotherapy"),
      "Yes", "No"
    ),
    index = row_number()
  )

n_samples <- 260
n_groups <- length(signature_groups)

base_latent <- matrix(rnorm(n_samples * n_groups), ncol = n_groups)
colnames(base_latent) <- names(signature_groups)

# Introduce interpretable cross-program relationships.
latent <- base_latent
latent[, "DNA repair"] <- 0.36 * latent[, "AR/SR signaling"] +
  sqrt(1 - 0.36^2) * base_latent[, "DNA repair"]
latent[, "Immune / microenvironment"] <- -0.48 * latent[, "AR/SR signaling"] +
  sqrt(1 - 0.48^2) * base_latent[, "Immune / microenvironment"]
latent[, "Lineage / plasticity"] <- -0.38 * latent[, "AR/SR signaling"] +
  0.30 * latent[, "Immune / microenvironment"] +
  0.82 * base_latent[, "Lineage / plasticity"]
latent[, "Oncogenic pathways"] <- 0.34 * latent[, "DNA repair"] +
  0.28 * latent[, "Lineage / plasticity"] +
  0.84 * base_latent[, "Oncogenic pathways"]
latent[, "Metabolism / prognosis"] <- 0.50 * latent[, "Oncogenic pathways"] -
  0.30 * latent[, "Immune / microenvironment"] +
  0.76 * base_latent[, "Metabolism / prognosis"]

score_matrix <- matrix(
  NA_real_, nrow = n_samples, ncol = nrow(signature_meta),
  dimnames = list(sprintf("Sample_%03d", seq_len(n_samples)), signature_meta$signature)
)

for (j in seq_len(nrow(signature_meta))) {
  group_name <- signature_meta$biological_category[j]
  primary <- latent[, group_name]
  secondary_group <- names(signature_groups)[1 + (match(group_name, names(signature_groups)) %% n_groups)]
  secondary <- latent[, secondary_group]
  loading <- runif(1, 0.68, 0.94)
  secondary_loading <- runif(1, -0.22, 0.22)
  score_matrix[, j] <- loading * primary +
    secondary_loading * secondary +
    rnorm(n_samples, sd = runif(1, 0.38, 0.62))
}

PSA <- 0.78 * latent[, "AR/SR signaling"] -
  0.25 * latent[, "Lineage / plasticity"] + rnorm(n_samples, sd = 0.55)
Ki67 <- 0.70 * latent[, "Metabolism / prognosis"] +
  0.45 * latent[, "Oncogenic pathways"] + rnorm(n_samples, sd = 0.55)

score_data <- as_tibble(score_matrix, rownames = "sample_id") %>%
  mutate(PSA = PSA, Ki67 = Ki67)

write.csv(
  score_data,
  "synthetic_signature_scores.csv",
  row.names = FALSE
)
write.csv(
  signature_meta,
  "signature_metadata.csv",
  row.names = FALSE
)
```

## Calculate Spearman correlations

```{r correlations}
signature_names <- signature_meta$signature
n_signatures <- length(signature_names)

corr_matrix <- cor(score_matrix, method = "spearman")
aux_matrix <- cor(
  score_matrix,
  cbind(PSA = PSA, Ki67 = Ki67),
  method = "spearman"
)

write.csv(
  tibble(signature = rownames(corr_matrix)) %>%
    bind_cols(as_tibble(corr_matrix)),
  "spearman_correlation_matrix.csv",
  row.names = FALSE
)

matrix_long <- as.data.frame(as.table(corr_matrix)) %>%
  transmute(
    row_signature = as.character(Var1),
    column_signature = as.character(Var2),
    correlation = Freq,
    row_index = match(row_signature, signature_names),
    x = match(column_signature, signature_names),
    y = n_signatures - row_index + 1
  )

aux_long <- as.data.frame(as.table(aux_matrix)) %>%
  transmute(
    signature = as.character(Var1),
    marker = as.character(Var2),
    correlation = Freq,
    signature_index = match(signature, signature_names),
    y = n_signatures - signature_index + 1,
    x = n_signatures + if_else(marker == "PSA", 1.7, 2.8)
  )

aux_bottom <- aux_long %>%
  transmute(
    x = signature_index,
    y = if_else(marker == "PSA", -1.7, -2.8),
    marker,
    correlation
  )
```

## Draw the annotated bubble heatmap

```{r plot, fig.width=14, fig.height=14, out.width="100%"}
category_palette <- c(
  "AR/SR signaling" = "#1F78B4",
  "DNA repair" = "#6A3D9A",
  "Immune / microenvironment" = "#33A02C",
  "Lineage / plasticity" = "#E31A1C",
  "Oncogenic pathways" = "#FF7F00",
  "Metabolism / prognosis" = "#B15928"
)

origin_palette <- c(
  "Prostate cancer" = "#E76F51",
  "Pan cancer" = "#457B9D",
  "CRPC" = "#8D5A97",
  "Breast" = "#F4A261",
  "Melanoma" = "#2A9D8F"
)

cycle_palette <- c("No" = "#E8EEF2", "Yes" = "#00A6A6")

annotation <- signature_meta %>%
  mutate(
    x = index,
    y = n_signatures - index + 1,
    category_colour = unname(category_palette[biological_category]),
    origin_colour = unname(origin_palette[origin]),
    cycle_colour = unname(cycle_palette[cell_cycle_enriched])
  )

top_category <- annotation %>% transmute(x, y = n_signatures + 2.0, fill = category_colour)
top_origin <- annotation %>% transmute(x, y = n_signatures + 3.0, fill = origin_colour)
top_cycle <- annotation %>% transmute(x, y = n_signatures + 4.0, fill = cycle_colour)

left_category <- annotation %>% transmute(x = -1.0, y, fill = category_colour)
left_origin <- annotation %>% transmute(x = -2.0, y, fill = origin_colour)
left_cycle <- annotation %>% transmute(x = -3.0, y, fill = cycle_colour)

matrix_grid <- expand_grid(x = seq_len(n_signatures), y = seq_len(n_signatures))
right_grid <- expand_grid(
  x = c(n_signatures + 1.7, n_signatures + 2.8),
  y = seq_len(n_signatures)
)
bottom_grid <- expand_grid(
  x = seq_len(n_signatures),
  y = c(-1.7, -2.8)
)

main_plot <- ggplot() +
  geom_tile(
    data = matrix_grid,
    aes(x, y), width = 0.92, height = 0.92,
    fill = "#FCFCFC", colour = "#D8DEE3", linewidth = 0.22
  ) +
  geom_tile(
    data = right_grid,
    aes(x, y), width = 0.86, height = 0.92,
    fill = "#FCFCFC", colour = "#D8DEE3", linewidth = 0.22
  ) +
  geom_tile(
    data = bottom_grid,
    aes(x, y), width = 0.92, height = 0.86,
    fill = "#FCFCFC", colour = "#D8DEE3", linewidth = 0.22
  ) +
  geom_point(
    data = matrix_long,
    aes(x, y, colour = correlation, size = abs(correlation)),
    alpha = 0.96
  ) +
  geom_point(
    data = aux_long,
    aes(x, y, colour = correlation, size = abs(correlation)),
    alpha = 0.96
  ) +
  geom_point(
    data = aux_bottom,
    aes(x, y, colour = correlation, size = abs(correlation)),
    alpha = 0.96
  ) +
  geom_tile(data = top_category, aes(x, y), fill = top_category$fill,
            width = 0.82, height = 0.75, colour = "white", linewidth = 0.2) +
  geom_tile(data = top_origin, aes(x, y), fill = top_origin$fill,
            width = 0.82, height = 0.75, colour = "white", linewidth = 0.2) +
  geom_tile(data = top_cycle, aes(x, y), fill = top_cycle$fill,
            width = 0.82, height = 0.75, colour = "white", linewidth = 0.2) +
  geom_tile(data = left_category, aes(x, y), fill = left_category$fill,
            width = 0.75, height = 0.82, colour = "white", linewidth = 0.2) +
  geom_tile(data = left_origin, aes(x, y), fill = left_origin$fill,
            width = 0.75, height = 0.82, colour = "white", linewidth = 0.2) +
  geom_tile(data = left_cycle, aes(x, y), fill = left_cycle$fill,
            width = 0.75, height = 0.82, colour = "white", linewidth = 0.2) +
  geom_text(
    data = tibble(
      x = seq_len(n_signatures),
      y = n_signatures + 0.62,
      label = signature_names
    ),
    aes(x, y, label = label), angle = 90, hjust = 0, vjust = 0.5,
    size = 2.35, colour = "#1F2933"
  ) +
  geom_text(
    data = tibble(
      x = -3.75,
      y = n_signatures - seq_len(n_signatures) + 1,
      label = signature_names
    ),
    aes(x, y, label = label), hjust = 1,
    size = 2.45, colour = "#1F2933"
  ) +
  annotate(
    "text", x = c(-3.75, -3.75, -3.75),
    y = n_signatures + c(2, 3, 4),
    label = c("Biological category", "Origin", "Cell-cycle enriched"),
    hjust = 1, size = 2.7, fontface = "bold", colour = "#39434A"
  ) +
  annotate(
    "text", x = c(-1, -2, -3), y = n_signatures + 0.62,
    label = c("Category", "Origin", "Cycle"),
    angle = 90, hjust = 0, size = 2.45, fontface = "bold", colour = "#39434A"
  ) +
  annotate(
    "text", x = c(n_signatures + 1.7, n_signatures + 2.8),
    y = n_signatures + 0.62, label = c("PSA", "Ki67"),
    angle = 90, hjust = 0, size = 2.7, fontface = "bold"
  ) +
  annotate(
    "text", x = 0.20, y = c(-1.7, -2.8),
    label = c("PSA", "Ki67"), hjust = 1,
    size = 2.7, fontface = "bold"
  ) +
  annotate(
    "rect", xmin = -9.25, xmax = -6.75,
    ymin = n_signatures + 1.55, ymax = n_signatures + 4.45,
    fill = "white", colour = "#5E6970", linewidth = 0.45
  ) +
  annotate(
    "segment", x = -9.25, xend = -6.75,
    y = n_signatures + 4.45, yend = n_signatures + 1.55,
    colour = "#5E6970", linewidth = 0.45
  ) +
  annotate(
    "text", x = -8.50, y = n_signatures + 3.90,
    label = "Shared\nprogram", size = 2.35, lineheight = 0.9
  ) +
  annotate(
    "text", x = -7.50, y = n_signatures + 2.05,
    label = "Context\nspecific", size = 2.35, lineheight = 0.9
  ) +
  annotate(
    "text", x = -8.60, y = n_signatures + 1.75,
    label = "Lower triangle", size = 2.2, angle = 45
  ) +
  annotate(
    "text", x = -7.35, y = n_signatures + 4.15,
    label = "Upper triangle", size = 2.2, angle = 45
  ) +
  scale_colour_gradient2(
    low = "#B2182B", mid = "#F5F5F5", high = "#2166AC",
    midpoint = 0, limits = c(-1, 1),
    breaks = c(-1, -0.5, 0, 0.5, 1),
    name = "Spearman correlation",
    guide = guide_colourbar(
      title.position = "top", title.hjust = 0.5,
      barwidth = grid::unit(52, "mm"), barheight = grid::unit(3.8, "mm")
    )
  ) +
  scale_size_area(max_size = 5.0, limits = c(0, 1), guide = "none") +
  coord_fixed(
    xlim = c(-9.45, n_signatures + 3.5),
    ylim = c(-3.45, n_signatures + 4.75),
    clip = "off",
    expand = FALSE
  ) +
  labs(
    title = "Cell-state signature correlation landscape",
    subtitle = "Circle area encodes |Spearman rho|; blue indicates positive and red indicates negative correlation"
  ) +
  theme_void(base_size = 10) +
  theme(
    plot.title = element_text(face = "bold", size = 17, hjust = 0.5, colour = "#20262B"),
    plot.subtitle = element_text(size = 9.5, hjust = 0.5, colour = "#59636B", margin = margin(b = 8)),
    plot.margin = margin(8, 10, 2, 10),
    legend.position = "bottom",
    legend.title = element_text(size = 8.5, face = "bold"),
    legend.text = element_text(size = 7.5)
  )

legend_panel <- function(values, colours, title, ncol = 1) {
  labels <- names(values)
  n <- length(labels)
  rows <- ceiling(n / ncol)
  dat <- tibble(
    label = labels,
    colour = unname(colours[labels]),
    column = rep(seq_len(ncol), each = rows, length.out = n),
    row = rep(seq(rows, 1), times = ncol, length.out = n)
  )

  ggplot(dat) +
    geom_point(
      aes(x = column * 3.3, y = row),
      shape = 22, size = 4.1, stroke = 0.35,
      fill = dat$colour, colour = "#667078"
    ) +
    geom_text(
      aes(x = column * 3.3 + 0.30, y = row, label = label),
      hjust = 0, size = 2.85, colour = "#283038"
    ) +
    annotate(
      "text", x = 3.3, y = rows + 0.85,
      label = title, hjust = 0, fontface = "bold", size = 3.1
    ) +
    coord_cartesian(
      xlim = c(3.0, ncol * 3.3 + 3.0),
      ylim = c(0.5, rows + 1.2), clip = "off"
    ) +
    theme_void() +
    theme(plot.margin = margin(0, 4, 0, 4))
}

category_legend <- legend_panel(
  setNames(names(category_palette), names(category_palette)),
  category_palette, "Biological category", ncol = 2
)
origin_legend <- legend_panel(
  setNames(names(origin_palette), names(origin_palette)),
  origin_palette, "Origin of signature", ncol = 2
)
cycle_legend <- legend_panel(
  setNames(names(cycle_palette), names(cycle_palette)),
  cycle_palette, "Cell-cycle enriched", ncol = 1
)

metadata_legends <- category_legend + origin_legend + cycle_legend +
  plot_layout(widths = c(1.35, 1.1, 0.65))

final_plot <- main_plot / metadata_legends +
  plot_layout(heights = c(12.2, 2.0))

ggsave(
  "correlation_bubble_heatmap.png",
  plot = final_plot,
  width = 14, height = 14, dpi = 320, bg = "white"
)

final_plot
```

## Output files

- `correlation_bubble_heatmap.png`: high-resolution figure.
- `synthetic_signature_scores.csv`: synthetic sample-by-signature score table.
- `signature_metadata.csv`: signature annotations.
- `spearman_correlation_matrix.csv`: complete signature correlation matrix.
