---
title: "AUC and Mean-Difference Estimation Plot"
author: "Synthetic example"
output:
  html_document:
    self_contained: true
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, warning = FALSE, message = FALSE,
  fig.align = "center", dpi = 160
)
library(ggplot2)
library(dplyr)
library(tidyr)
library(ggbeeswarm)
library(patchwork)
set.seed(1031)
```

## Synthetic paired AUC data

Three paired comparisons are simulated with 125 observations per group. The right panel displays bootstrap distributions of the paired mean differences.

```{r data}
n_subjects <- 125

comparison_1 <- tibble(
  subject = seq_len(n_subjects),
  Solvent = rnorm(n_subjects, mean = 0.0, sd = 6.0)
) %>%
  mutate(Air = Solvent + rnorm(n_subjects, mean = 1.1, sd = 4.2))

comparison_2 <- tibble(
  subject = seq_len(n_subjects),
  Solvent = rnorm(n_subjects, mean = 0.0, sd = 6.2)
) %>%
  mutate(IPM = Solvent + rnorm(n_subjects, mean = 19.5, sd = 11.0))

comparison_3 <- tibble(
  subject = seq_len(n_subjects),
  Air = rnorm(n_subjects, mean = 1.2, sd = 6.0)
) %>%
  mutate(IPM = Air + rnorm(n_subjects, mean = 17.5, sd = 10.5))

left_data <- bind_rows(
  comparison_1 %>%
    pivot_longer(c(Solvent, Air), names_to = "group", values_to = "AUC") %>%
    mutate(comparison = "Air - Solvent", row_y = if_else(group == "Solvent", 6, 5)),
  comparison_2 %>%
    pivot_longer(c(Solvent, IPM), names_to = "group", values_to = "AUC") %>%
    mutate(comparison = "IPM - Solvent", row_y = if_else(group == "Solvent", 4, 3)),
  comparison_3 %>%
    pivot_longer(c(Air, IPM), names_to = "group", values_to = "AUC") %>%
    mutate(comparison = "IPM - Air", row_y = if_else(group == "Air", 2, 1))
) %>%
  mutate(
    group = factor(group, levels = c("Solvent", "Air", "IPM")),
    AUC = pmin(60, pmax(-12, AUC))
  )

paired_differences <- bind_rows(
  comparison_1 %>% transmute(comparison = "Air - Solvent", difference = Air - Solvent),
  comparison_2 %>% transmute(comparison = "IPM - Solvent", difference = IPM - Solvent),
  comparison_3 %>% transmute(comparison = "IPM - Air", difference = IPM - Air)
)

bootstrap_mean <- function(x, n_boot = 4000) {
  replicate(n_boot, mean(sample(x, length(x), replace = TRUE)))
}

bootstrap_data <- paired_differences %>%
  group_by(comparison) %>%
  summarise(
    bootstrap_mean = list(bootstrap_mean(difference)),
    .groups = "drop"
  ) %>%
  unnest_longer(bootstrap_mean) %>%
  mutate(
    y = recode(
      comparison,
      `Air - Solvent` = 5,
      `IPM - Solvent` = 3,
      `IPM - Air` = 1
    ),
    density_group = if_else(comparison == "Air - Solvent", "Air", "IPM")
  )

difference_summary <- paired_differences %>%
  group_by(comparison) %>%
  summarise(
    mean_difference = mean(difference),
    ci_low = quantile(bootstrap_mean(difference), 0.025),
    ci_high = quantile(bootstrap_mean(difference), 0.975),
    .groups = "drop"
  ) %>%
  mutate(
    y = recode(
      comparison,
      `Air - Solvent` = 5,
      `IPM - Solvent` = 3,
      `IPM - Air` = 1
    ),
    density_group = if_else(comparison == "Air - Solvent", "Air", "IPM")
  )

write.csv(left_data, "synthetic_auc_observations.csv", row.names = FALSE)
write.csv(bootstrap_data, "bootstrap_mean_differences.csv", row.names = FALSE)
write.csv(difference_summary, "mean_difference_summary.csv", row.names = FALSE)
```

## Estimation plot

```{r plot, fig.width=7.6, fig.height=6.4, out.width="82%"}
group_colours <- c(
  "Solvent" = "#FC8D62",
  "Air" = "#8DA0CB",
  "IPM" = "#66C2A5"
)

left_panel <- ggplot(left_data, aes(AUC, row_y)) +
  ggbeeswarm::geom_quasirandom(
    aes(colour = group, group = interaction(comparison, group)),
    groupOnX = FALSE,
    width = 0.22,
    shape = 21,
    fill = "white",
    size = 1.65,
    stroke = 0.55,
    alpha = 0.98,
    show.legend = FALSE
  ) +
  scale_colour_manual(values = group_colours) +
  annotate(
    "text",
    x = -15.5, y = c(6, 5, 4, 3, 2, 1),
    label = c("Solvent", "Air", "Solvent", "IPM", "Air", "IPM"),
    colour = unname(group_colours[c("Solvent", "Air", "Solvent", "IPM", "Air", "IPM")]),
    hjust = 1, size = 3.5
  ) +
  annotate(
    "segment",
    x = -12.5, xend = -12.5,
    y = c(5, 3, 1), yend = c(6, 4, 2),
    linewidth = 0.45, colour = "#555555"
  ) +
  annotate(
    "segment",
    x = -12.5, xend = -10.0,
    y = c(6, 5, 4, 3, 2, 1), yend = c(6, 5, 4, 3, 2, 1),
    linewidth = 0.45, colour = "#555555"
  ) +
  annotate(
    "text", x = 52, y = c(5.5, 3.5, 1.5),
    label = "italic(n) == 125", parse = TRUE,
    hjust = 0.5, size = 3.3
  ) +
  annotate(
    "text", x = -18.5, y = 6.62,
    label = "f", fontface = "bold", size = 5.6
  ) +
  scale_x_continuous(
    limits = c(-19, 62), breaks = c(0, 50),
    expand = c(0, 0)
  ) +
  scale_y_continuous(limits = c(0.45, 6.75), expand = c(0, 0)) +
  labs(x = "AUC", y = NULL) +
  coord_cartesian(clip = "off") +
  theme_classic(base_size = 12) +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    axis.title.x = element_text(size = 13, margin = margin(t = 8)),
    axis.text.x = element_text(size = 11, colour = "black"),
    axis.line.x = element_line(linewidth = 0.6),
    axis.ticks.x = element_line(linewidth = 0.55),
    axis.ticks.length = grid::unit(2.0, "mm"),
    plot.margin = margin(10, 2, 8, 15)
  )

right_panel <- ggplot(
  bootstrap_data,
  aes(x = bootstrap_mean, y = y, group = comparison, fill = density_group)
) +
  geom_vline(xintercept = 0, colour = "#595959", linewidth = 0.55) +
  geom_violin(
    orientation = "y", width = 0.48,
    scale = "width", trim = TRUE,
    colour = NA, alpha = 0.74
  ) +
  geom_errorbarh(
    data = difference_summary,
    aes(
      y = y,
      xmin = ci_low,
      xmax = ci_high
    ),
    inherit.aes = FALSE,
    height = 0,
    linewidth = 0.75,
    colour = "black"
  ) +
  geom_point(
    data = difference_summary,
    aes(x = mean_difference, y = y),
    inherit.aes = FALSE,
    size = 2.3, colour = "black"
  ) +
  geom_point(
    data = tibble(x = 0, y = c(6, 4, 2)),
    aes(x, y), inherit.aes = FALSE,
    size = 2.2, colour = "black"
  ) +
  annotate(
    "text", x = 27.0, y = c(3, 1),
    label = "#", size = 4.0, hjust = 0.5
  ) +
  scale_fill_manual(values = group_colours[c("Air", "IPM")], guide = "none") +
  scale_x_continuous(
    limits = c(-2.5, 29), breaks = c(0, 20),
    expand = c(0, 0)
  ) +
  scale_y_continuous(limits = c(0.45, 6.75), expand = c(0, 0)) +
  labs(x = "Mean difference", y = NULL) +
  coord_cartesian(clip = "off") +
  theme_classic(base_size = 12) +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    axis.title.x = element_text(size = 13, margin = margin(t = 8)),
    axis.text.x = element_text(size = 11, colour = "black"),
    axis.line.x = element_line(linewidth = 0.6),
    axis.ticks.x = element_line(linewidth = 0.55),
    axis.ticks.length = grid::unit(2.0, "mm"),
    plot.margin = margin(10, 10, 8, 2)
  )

final_plot <- left_panel + right_panel +
  plot_layout(widths = c(1.72, 1.0))

ggsave(
  "auc_mean_difference_estimation_plot.png",
  final_plot,
  width = 7.6, height = 6.4, dpi = 320, bg = "white"
)

final_plot
```

## Files

- `auc_mean_difference_estimation_plot.png`: high-resolution combined figure.
- `synthetic_auc_observations.csv`: paired AUC observations.
- `bootstrap_mean_differences.csv`: bootstrap sampling distributions.
- `mean_difference_summary.csv`: mean differences and confidence intervals.
