---
title: "Global Wetland Cover Loss"
author: "Synthetic example"
output:
  html_document:
    self_contained: true
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, warning = FALSE, message = FALSE,
  fig.align = "center", dpi = 160
)
library(ggplot2)
library(dplyr)
library(tidyr)
library(sf)
library(rnaturalearth)
set.seed(1031)
```

## Synthetic spatial data

This reproduction uses a regular longitude–latitude grid and synthetic wetland-loss values. It is intended to reproduce the visual structure of the reference figure, not to provide historical estimates.

```{r simulate-data}
world <- ne_countries(scale = "medium", returnclass = "sf") %>%
  filter(continent != "Antarctica") %>%
  st_make_valid()

grid <- expand_grid(
  longitude = seq(-179.75, 179.75, by = 0.5),
  latitude = seq(-58.75, 82.75, by = 0.5)
)

# Create an organic, spatially autocorrelated texture on the complete grid.
# Combining fine- and medium-scale smoothed fields avoids both isolated square
# pixels and directional Fourier stripes.
n_lon <- length(unique(grid$longitude))
n_lat <- length(unique(grid$latitude))

smooth_field <- function(z, kernel) {
  kernel <- kernel / sum(kernel)
  by_column <- apply(
    z, 2,
    function(v) as.numeric(stats::filter(v, kernel, sides = 2, circular = TRUE))
  )
  t(apply(
    by_column, 1,
    function(v) as.numeric(stats::filter(v, kernel, sides = 2, circular = TRUE))
  ))
}

raw_noise <- matrix(rnorm(n_lon * n_lat), nrow = n_lat, ncol = n_lon)
fine_field <- smooth_field(raw_noise, c(1, 2, 1))
medium_field <- smooth_field(raw_noise, c(1, 2, 3, 2, 1))
texture_raw <- 0.68 * fine_field + 0.32 * medium_field
texture_limits <- quantile(texture_raw, c(0.02, 0.98))
grid$spatial_texture <- pmin(
  1, pmax(0, (as.vector(texture_raw) - texture_limits[1]) /
              diff(texture_limits))
)

grid_sf <- st_as_sf(
  grid, coords = c("longitude", "latitude"),
  crs = 4326, remove = FALSE
)

on_land <- lengths(st_intersects(grid_sf, world)) > 0
land_grid <- grid[on_land, ]

spatial_kernel <- function(lon, lat, centre_lon, centre_lat,
                           lon_sd, lat_sd) {
  # Wrap longitudinal distance around the date line.
  dlon <- abs(lon - centre_lon)
  dlon <- pmin(dlon, 360 - dlon)
  exp(-0.5 * ((dlon / lon_sd)^2 + ((lat - centre_lat) / lat_sd)^2))
}

wetland_centres <- tribble(
  ~lon, ~lat, ~lon_sd, ~lat_sd, ~weight,
  # North America: eastern interior, Great Lakes, and Pacific coast
   -84,   37,      24,      12,    1.10,
   -91,   46,      18,       9,    0.70,
  -121,   39,       8,      17,    0.72,
  -102,   24,      18,      10,    0.52,
  # South America: Amazon margins, Andes, Pantanal, and Atlantic coast
   -61,   -5,      23,       8,    0.30,
   -74,  -12,       6,      23,    0.35,
   -57,  -19,      12,       8,    0.72,
   -49,  -27,      15,      13,    0.86,
  # Europe and the broad Eurasian wetland belt
     8,   50,      23,      11,    1.10,
    18,   63,      16,       8,    0.55,
    47,   55,      32,       9,    0.38,
    87,   56,      38,      10,    0.30,
   126,   49,      21,      10,    0.38,
  # South, East, and Southeast Asia
    79,   23,      16,      12,    1.12,
   112,   32,      23,      14,    1.28,
   125,   44,      15,       9,    0.82,
   105,   12,      17,      12,    1.02,
   136,   37,       9,      10,    0.55,
  # African wetland belts and Madagascar
     2,   13,      37,       5,    0.48,
    30,   -2,      16,      15,    0.38,
    25,  -24,      14,      10,    0.40,
    47,  -20,       8,      10,    0.58,
  # Australian and New Zealand coastal regions
   149,  -28,       8,      15,    0.66,
   116,  -32,       7,       6,    0.40,
   173,  -41,       7,       8,    0.55
)

wetland_score <- rep(0, nrow(land_grid))
for (i in seq_len(nrow(wetland_centres))) {
  wetland_score <- wetland_score + wetland_centres$weight[i] *
    spatial_kernel(
      land_grid$longitude, land_grid$latitude,
      wetland_centres$lon[i], wetland_centres$lat[i],
      wetland_centres$lon_sd[i], wetland_centres$lat_sd[i]
    )
}

# Convert the regional kernels into a habitat-suitability surface.
habitat_suitability <- 1 - exp(-wetland_score)

# Habitat suitability controls how much of the local textured field is kept.
# Strong wetland regions remain dense but retain small holes; marginal regions
# break into sparse, irregular patches.
presence_threshold <- 0.84 - 0.48 * habitat_suitability
wetland_present <- land_grid$spatial_texture > presence_threshold &
  habitat_suitability > 0.11

pressure_centres <- tribble(
  ~lon, ~lat, ~lon_sd, ~lat_sd, ~weight,
   -84,   37,      22,      12,    1.10,
  -121,   37,       9,      14,    0.58,
   -49,  -26,      16,      13,    0.98,
   -72,   -9,       8,      22,    0.62,
     9,   50,      24,      12,    1.50,
    43,   53,      30,       9,    0.32,
   105,   55,      40,      10,    0.18,
    79,   23,      16,      12,    1.35,
   115,   32,      21,      13,    1.75,
   126,   45,      14,       9,    1.05,
   105,   12,      16,      11,    1.45,
     1,   13,      37,       5,    0.28,
    31,   -3,      16,      15,    0.30,
    26,  -25,      14,      10,    0.30,
    47,  -20,       8,       9,    0.52,
   149,  -30,       8,      14,    0.48,
   116,  -32,       7,       6,    0.38,
   173,  -41,       7,       8,    0.68
)

human_pressure <- rep(0, nrow(land_grid))
for (i in seq_len(nrow(pressure_centres))) {
  human_pressure <- human_pressure + pressure_centres$weight[i] *
    spatial_kernel(
      land_grid$longitude, land_grid$latitude,
      pressure_centres$lon[i], pressure_centres$lat[i],
      pressure_centres$lon_sd[i], pressure_centres$lat_sd[i]
    )
}

wetland_loss <- land_grid %>%
  mutate(
    wetland_present = wetland_present,
    wetland_score = wetland_score,
    human_pressure = 1 - exp(-human_pressure),
    spatial_texture = spatial_texture,
    percentage_lost = pmin(
      100,
      pmax(
        1,
        3 + 88 * human_pressure +
          13 * (spatial_texture - 0.5)
      )
    )
  ) %>%
  filter(wetland_present) %>%
  select(longitude, latitude, percentage_lost)

write.csv(wetland_loss, "synthetic_wetland_loss_grid.csv", row.names = FALSE)
write.csv(wetland_centres, "wetland_region_parameters.csv", row.names = FALSE)
write.csv(pressure_centres, "wetland_loss_pressure_parameters.csv", row.names = FALSE)
```

## Rectangular world map

```{r map, fig.width=13, fig.height=5.5, out.width="100%"}
legend_dummy <- tibble(
  longitude = -179,
  latitude = -59,
  status = "<1% loss or\n<1% wetland\nin 1700"
)

wetland_map <- ggplot() +
  geom_sf(
    data = world,
    fill = "#E5E6E6", colour = "#C6C9C9", linewidth = 0.18
  ) +
  geom_tile(
    data = wetland_loss,
    aes(longitude, latitude, fill = percentage_lost),
    width = 0.52, height = 0.52, alpha = 0.97
  ) +
  geom_sf(
    data = world,
    fill = NA, colour = "#D0D2D2", linewidth = 0.14
  ) +
  geom_point(
    data = legend_dummy,
    aes(longitude, latitude, colour = status),
    shape = 15, size = 4.5, alpha = 0
  ) +
  scale_fill_gradientn(
    colours = c("#FFE45B", "#FFC22E", "#FF8A18", "#F04A0B", "#CC1707", "#8D0000"),
    limits = c(1, 100),
    breaks = c(1, 25, 50, 75, 100),
    name = "Percentage of wetland\ncover lost (% of wetland\narea in 1700)",
    guide = guide_colourbar(
      order = 2,
      title.position = "top", title.hjust = 0.5,
      barheight = grid::unit(49, "mm"),
      barwidth = grid::unit(5.5, "mm"),
      ticks.colour = "#333333", frame.colour = "#666666"
    )
  ) +
  scale_colour_manual(
    values = c("<1% loss or\n<1% wetland\nin 1700" = "#D7D9D9"),
    name = NULL,
    guide = guide_legend(
      order = 1,
      override.aes = list(alpha = 1, size = 5.5, shape = 15)
    )
  ) +
  coord_sf(
    xlim = c(-180, 180), ylim = c(-60, 84),
    expand = FALSE, datum = NA
  ) +
  theme_void(base_size = 11) +
  theme(
    panel.background = element_rect(fill = "white", colour = "#505050", linewidth = 0.65),
    plot.background = element_rect(fill = "white", colour = NA),
    plot.margin = margin(10, 10, 10, 10),
    legend.position = "right",
    legend.box = "vertical",
    legend.box.just = "left",
    legend.spacing.y = grid::unit(3, "mm"),
    legend.title = element_text(size = 10.5, face = "bold", lineheight = 0.95),
    legend.text = element_text(size = 9, lineheight = 0.95),
    legend.key = element_rect(fill = "white", colour = NA)
  )

ggsave(
  "wetland_cover_loss_map.png",
  wetland_map,
  width = 13, height = 5.5, dpi = 320, bg = "white"
)

wetland_map
```

## Files

- `wetland_cover_loss_map.png`: high-resolution rectangular map.
- `synthetic_wetland_loss_grid.csv`: synthetic gridded percentage-loss data.
- `wetland_region_parameters.csv`: spatial-kernel parameters.
- `wetland_loss_pressure_parameters.csv`: regional loss-intensity parameters.
