---
title: "Global Wetland Cover Loss"
author: "Synthetic example"
output:
  html_document:
    self_contained: true
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, warning = FALSE, message = FALSE,
  fig.align = "center", dpi = 160
)
library(ggplot2)
library(dplyr)
library(tidyr)
library(sf)
library(rnaturalearth)
set.seed(1031)
```

## Synthetic spatial data

This reproduction uses a regular longitude–latitude grid and synthetic wetland-loss values. It is intended to reproduce the visual structure of the reference figure, not to provide historical estimates.

```{r simulate-data}
world <- ne_countries(scale = "medium", returnclass = "sf") %>%
  filter(continent != "Antarctica") %>%
  st_make_valid()

grid <- expand_grid(
  longitude = seq(-179.5, 179.5, by = 1.0),
  latitude = seq(-58.5, 82.5, by = 1.0)
)

grid_sf <- st_as_sf(
  grid, coords = c("longitude", "latitude"),
  crs = 4326, remove = FALSE
)

on_land <- lengths(st_intersects(grid_sf, world)) > 0
land_grid <- grid[on_land, ]

spatial_kernel <- function(lon, lat, centre_lon, centre_lat,
                           lon_sd, lat_sd) {
  # Wrap longitudinal distance around the date line.
  dlon <- abs(lon - centre_lon)
  dlon <- pmin(dlon, 360 - dlon)
  exp(-0.5 * ((dlon / lon_sd)^2 + ((lat - centre_lat) / lat_sd)^2))
}

wetland_centres <- tribble(
  ~lon, ~lat, ~lon_sd, ~lat_sd, ~weight,
   -92,   34,      24,      13,    0.95,
   -61,   -5,      25,      13,    0.82,
   -57,  -20,      13,       8,    0.72,
    10,   50,      28,      12,    1.00,
    45,   55,      38,      15,    0.78,
    87,   55,      40,      15,    0.67,
   114,   31,      23,      13,    1.00,
    80,   23,      18,      12,    0.95,
   105,   12,      18,      12,    0.93,
    25,    0,      24,      14,    0.66,
    47,  -20,      10,       9,    0.63,
   145,  -27,      18,      13,    0.75
)

wetland_score <- rep(0, nrow(land_grid))
for (i in seq_len(nrow(wetland_centres))) {
  wetland_score <- wetland_score + wetland_centres$weight[i] *
    spatial_kernel(
      land_grid$longitude, land_grid$latitude,
      wetland_centres$lon[i], wetland_centres$lat[i],
      wetland_centres$lon_sd[i], wetland_centres$lat_sd[i]
    )
}

# Wetland occurrence is spatially clustered and sparse outside the main regions.
occurrence_probability <- pmin(
  0.94,
  0.035 + 0.74 * (1 - exp(-1.35 * wetland_score)) +
    0.08 * exp(-(land_grid$latitude / 28)^2)
)

wetland_present <- runif(nrow(land_grid)) < occurrence_probability

pressure_centres <- tribble(
  ~lon, ~lat, ~lon_sd, ~lat_sd, ~weight,
   -88,   36,      20,      12,    0.72,
    10,   50,      25,      12,    0.88,
    79,   23,      18,      12,    0.84,
   116,   32,      20,      12,    1.00,
   106,   12,      16,      11,    0.88,
   -58,  -22,      17,      11,    0.62,
    30,   -2,      21,      14,    0.48,
   146,  -31,      16,      11,    0.58
)

human_pressure <- rep(0, nrow(land_grid))
for (i in seq_len(nrow(pressure_centres))) {
  human_pressure <- human_pressure + pressure_centres$weight[i] *
    spatial_kernel(
      land_grid$longitude, land_grid$latitude,
      pressure_centres$lon[i], pressure_centres$lat[i],
      pressure_centres$lon_sd[i], pressure_centres$lat_sd[i]
    )
}

wetland_loss <- land_grid %>%
  mutate(
    wetland_present = wetland_present,
    wetland_score = wetland_score,
    human_pressure = human_pressure,
    percentage_lost = pmin(
      100,
      pmax(1, 8 + 68 * (1 - exp(-1.15 * human_pressure)) +
             17 * runif(n()) + rnorm(n(), sd = 4))
    )
  ) %>%
  filter(wetland_present) %>%
  select(longitude, latitude, percentage_lost)

write.csv(wetland_loss, "synthetic_wetland_loss_grid.csv", row.names = FALSE)
write.csv(wetland_centres, "wetland_region_parameters.csv", row.names = FALSE)
```

## Rectangular world map

```{r map, fig.width=13, fig.height=5.5, out.width="100%"}
legend_dummy <- tibble(
  longitude = -179,
  latitude = -59,
  status = "<1% wetland in 1700\nor unavailable"
)

wetland_map <- ggplot() +
  geom_sf(
    data = world,
    fill = "#E5E6E6", colour = "#C6C9C9", linewidth = 0.18
  ) +
  geom_tile(
    data = wetland_loss,
    aes(longitude, latitude, fill = percentage_lost),
    width = 1.02, height = 1.02, alpha = 0.96
  ) +
  geom_sf(
    data = world,
    fill = NA, colour = "#D0D2D2", linewidth = 0.14
  ) +
  geom_point(
    data = legend_dummy,
    aes(longitude, latitude, colour = status),
    shape = 15, size = 4.5, alpha = 0
  ) +
  scale_fill_gradientn(
    colours = c("#FFE45B", "#FFC22E", "#FF8A18", "#F04A0B", "#CC1707", "#8D0000"),
    limits = c(1, 100),
    breaks = c(1, 25, 50, 75, 100),
    name = "Percentage of wetland\ncover lost (% of wetland\narea in 1700)",
    guide = guide_colourbar(
      order = 2,
      title.position = "top", title.hjust = 0.5,
      barheight = grid::unit(49, "mm"),
      barwidth = grid::unit(5.5, "mm"),
      ticks.colour = "#333333", frame.colour = "#666666"
    )
  ) +
  scale_colour_manual(
    values = c("<1% wetland in 1700\nor unavailable" = "#D7D9D9"),
    name = NULL,
    guide = guide_legend(
      order = 1,
      override.aes = list(alpha = 1, size = 5.5, shape = 15)
    )
  ) +
  coord_sf(
    xlim = c(-180, 180), ylim = c(-60, 84),
    expand = FALSE, datum = NA
  ) +
  theme_void(base_size = 11) +
  theme(
    panel.background = element_rect(fill = "white", colour = "#505050", linewidth = 0.65),
    plot.background = element_rect(fill = "white", colour = NA),
    plot.margin = margin(10, 10, 10, 10),
    legend.position = "right",
    legend.box = "vertical",
    legend.box.just = "left",
    legend.spacing.y = grid::unit(3, "mm"),
    legend.title = element_text(size = 10.5, face = "bold", lineheight = 0.95),
    legend.text = element_text(size = 9, lineheight = 0.95),
    legend.key = element_rect(fill = "white", colour = NA)
  )

ggsave(
  "wetland_cover_loss_map.png",
  wetland_map,
  width = 13, height = 5.5, dpi = 320, bg = "white"
)

wetland_map
```

## Files

- `wetland_cover_loss_map.png`: high-resolution rectangular map.
- `synthetic_wetland_loss_grid.csv`: synthetic gridded percentage-loss data.
- `wetland_region_parameters.csv`: spatial-kernel parameters.
